Compararison of analytical and artificial neural network models in predicting hot deformation behavior of AISI P20 and interpolation--extrapolation capacity validation

### 2022-10-14 Article.tex
Remove scalebox in Latex
Update and clean of bibliography. Remove doubles

%--------------------------------------------------------------------------
\section{Introduction\label{sec:Introduction}}
%--------------------------------------------------------------------------
Continuous casting and rolling mechanisms are an integral part of the technological methods of medium deformation through a continuous water-cooled crystalliser before entering a rolling mill. Some of the advantages of this method are increased metal yield, energy saving, improved quality of continuous casting and easier mechanisation and automation . Deformation being a very complex process involving both elastic and plastic deformation, is accompanied by high temperatures, high strain rates and complex and variable friction conditions \cite{he2013comparative}. However, a quantitative study and analysis of the deformation mechanism does not provide accurate analytical solutions . Moreover, conducting an experiment is not only time-consuming but also requires a high level of funding \cite{changizian2012high}. To mitigate these two factors (to name but a few), the Finite Element Method (FEM) is widely used in the analysis and optimisation of process parameters of various materials . FEM effectively compensates for the shortcomings of the analytical and experimental methods and plays an important role in the correct formulation of technological parameters, the improvement of product quality and the in-depth knowledge of the deformation mechanism. For example FEM has been widely used to simulate and analyse steel hot forming processes, such as forging, rolling, extrusion and casting but the accuracy of the numerical solution provided by FEM for this type of analysis strongly depends  on the accuracy of the constitutive equation of the concerned material \cite{qin2010modeling, mandal2009constitutive, ji2018constitutive}.

Indeed, hot forming processes are an essential step in the manufacturing of mechanical components and directly affect the ﬁnal quality of the products such as microstructure and mechanical properties \cite{ashtiani2012constitutive}. Several metallurgical phenomena such as work hardening and dynamic softening (dynamic recovery and recrystallisation), which very often occur during hot working, influence the quality of the deformation of materials. The complexity of the deformation behavior, however, leads to the fact that the deformation can only be calculated accurately by using a computer code to model the response of the material under the specified loading conditions. To this end, a mathematical formulation is often used to represent the flow behavior of materials in a form that can be used in finite element software \cite{lin2008prediction,wu2012flow}.

Based on a thorough search of the literature many researchers have tried to predict the mechanical behavior of materials under hot working conditions using constitutive models. Therefore, various constitutive equations have been proposed to describe the flow behavior including phenomenological, physical and artiﬁcial neural network (ANN) models \cite{lin2011critical,shin2010phenomenological,rusinek2010thermo}. As for the semi-empirical models, they try to be at the border between the empirical and physical models. Among a large number of models, the Johnson--Cook (JC) \cite{Johnson-1983-Constitutive, Johnson-1985-Fracture} and the Zerilli--Armstrong (ZA) \cite{Zerilli-1987-Dislocation} models are the best known and available in finite element codes. The JC model is widely used because it is simple to use and has fewer constants to be determined \cite{Khan-2004-Quasi-static, NematNasser-2003-Thermomechanical}. However, this model, in its original state, contains some problems related to the phenomena describing the behavior of the material. We refer here to strain hardening, strain rate hardening, and thermal effects that are described separately in the JC model as analyzed by Jia \eal \cite{Jia-2021-Modified-JC}. To overcome this difficulty, several modified forms of the Johnson--Cook model have been developed in the past \cite{Rule-1998-Revised-JC, Vural-2003-Large, Lin-2010-Modified-JC, Lin-2012-Phenomenological, Li-2013-Modified-JC, Zhou-2019-Research, Zhang-2015-Modified-JC}. According to its more physical formulation, the ZA model is preferred over the JC model because it allows for the coupling between strain rate and temperature effects \cite{Johnson-1988-Evaluation, Voyiadjis-2005-Microstructural, Dey-2007-Influence}. However, the drawbacks of this model are related to the temperature limit value of the test. Indeed, this model applies only to low strain rates and for temperatures limited to $0.6~T_m$ (where $T_m$ is the melting temperature of the material) \cite{Chiou-2005-Strain, Lee-2005-Strain, Chen-2007-Comparative, Lee-2006-Effects}. Moreover, this model, in its original form, does not take into account the coupling between deformation and temperature effects \cite{Samantaray-2009-Thermo-viscoplastic}. In order to reduce these shortcomings (\eg applying this model over a wider temperature range) of the ZA model, Modified Zerilli--Armstrong forms (MZA) have been proposed by several authors over the last past years \cite{Nemat-2004-Plasticity, Lennon-2004-Influence, Muralli-2017-Performance, Cheng-2021-Modified-ZA, Muralli-2021-Finite}.
The hyperbolic sine Arrhenius-type constitutive model is one of the most widely used phenomenological models both for the prediction of the hot deformation behavior of materials and for the analysis of the microstructure of the studied material because   this model allows to estimate the deformation energy of the material \cite{jonas1969strength,zhang2012effect,mostafaei2012hot}. Its original formulation has been revised several times in order to properly describe the obtained experimental behavior. In order to increase the accuracy of the model, the effects of strain \cite{slooff2007constitutive, li2012flow, xu2013constitutive}, strain rate \cite{lin2008constitutive, mandal2009constitutive} have been considered by several researchers. For example, Sloof \eal \cite{slooff2007constitutive} introduced a strain-dependent parameter into the initial formulation of the Arhenius model so that their model could predict the high temperature flow behavior of a magnesium alloy. Lin \eal \cite{lin2008constitutive}, on the other hand, modified the model by introducing a strain rate compensation factor.

@@

%----------------------------------------------
\section{ Materials and Experiments \label{sec:Materials}}
%----------------------------------------------
%----------------------------------------------
\subsection{Experimental procedure}
%----------------------------------------------
The material used in this study is the modified carbon alloy AISI P20 whose chemical composition is given in the Table \ref{tab:Composition}. Cylindrical specimens were machined with an initial diameter of $10$ mm and a height of $15$ mm. Hot compression tests were conducted on a Gleeble-3800 thermo-simulator (Figure \ref{fig:Gleeble3800}), at temperatures of $1050^\circ$C,  $1050^\circ$C,  $1100^\circ$C,  $1150^\circ$C,  $1200^\circ$C and  $1250^\circ$C with strain rates of $0.001\ \text{s}^{-1}$, $0.01\ \text{s}^{-1}$, $0.1\ \text{s}^{-1}$, $1.0\ \text{s}^{-1}$, $2.0\ \text{s}^{-1}$ and $5.0\ \text{s}^{-1}$. Thin tantalum sheets were used as a lubricant material at the contact surface of anvils and specimen to minimize the frictions during the tests. The samples were heated up to temperature $1260^\circ$C  and were held for $5$ min to eliminate thermal gradients. Then cooled to test temperature where the specimen is held $1$ minute before deformation. After compression the specimen is quencted rapidly for microstructure studies. The procedure of this test is illustrated in Figure \ref{fig:Heating} where $h_0$ and $R_0$ are the height and radius before compression; $h$, $R_M$, and $R_T$ are respectively the height, big radius and small radius of the specimen after compression.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]
{Figures/Gleeble-3800}
\caption{Gleeble 3800 simulator system}
\label{fig:Gleeble3800}
\end{figure}
\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]
{Figures/GleebleDesign}
\caption{Schematic diagram of the experimental process.}
\label{fig:Heating}
\end{figure}
\begin{table}[h!]
\centering{}
\caption{Chemical Composition of the AISI P20 Steel (Weight Percent)}
\scalebox{0.9}{
\begin{tabular}{lcccccccc}
\hline
C  &Mn & Mo   & Si   & Ni       &Cr& Cu & Other& Balance\\
\hline
 0.30        &       0.89      &0.52		   &	0.34     &	 0.68          &1.86& 0.17& Microalloying& Iron\\
\hline
\label{tab:Composition}
\end{tabular}}
\end{table}
%------------------------------------------------------------------------
\subsection{Flow stress behavior}
%------------------------------------------------------------------------
The flow stress curves obtained from the compression tests performed on Gleeble-3800 simulator for each test condition are shown in Figure \ref{fig:rawData}. From these curves it can be seen that the flow stress increases with increasing strain rate but decreases with increasing temperature. It should be noted that the strain influences the flow stress. Indeed, for all strain values below $0.2$, the stress increases with strain before reaching a peak value at low strain rates and then decreases to maintain a stable value until the end of the test. The peak stress value increases with strain rate but decreases with temperature. For higher strain rates ($1.0\ \text{s}^{-1}$, $2.0\ \text{s}^{-1}$ and $5.0\ \text{s}^{-1}$) the flow stress increases throughout the test. The slight increase in stress at low strain rates when the strain is large is due to friction between the specimen and the grip during the test. For low strain rates, lubrication is lost over time and therefore friction occurs. This results in a progressive increase in stress.

@@

The JC model as mentioned above, is one of the most widely used analytical models as it can be applied to several materials under different conditions of strain, strain rate and temperature . However, the formulation of this model does not take into account the simultaneous effect of strain, strain rate and temperature. Indeed, it is formulated by describing the effect of each physical parameter (strain, strain rate and temperature) separately as a factor in the mathematical expression of the model. Hence its inability to describe the phenomenon of temperature-induced softening. The equation that describes this model is given as follows \cite{johnson1985fracture}:

@@

The predicted and experimental values are compared, as shown in Figure \ref{fig:iCorrelationJC}. The JC model is not much suitable to describe the ﬂow behavior of AISI P20 within entire deformation condition, and predicted values show a monotonous downward trend at various strain rates. The deviation between predicted values and experimental values is large for the low strain rates and almost acceptable at high strain rates. All parameters of JC model are in Table \ref{tab: JCparameters}.
\begin{figure}[!ht]
\centering
\includegraphics[width=1.02\columnwidth]
{Figures/CompExpJC}
\caption{Comparison between the experimental and predicted flow stresses by JC model}
\label{fig:iCorrelationJC}
\end{figure}
\begin{table}[h!]
\centering{}
\caption{Parameters' constants of Johnson-Cook Model}
\scalebox{1.2}{
\begin{tabular}{lccccc}
\hline
&         &             &		   &		  &\\
Parameters&$A(\text{MPa})$    &$B(\text{MPa})$        & $n$     & $C$   & $m$       \\
&         &             &		   &	     &	    \\
\hline
Values&$43.0988$&$114.518$& $0.313969$&$0.0995438$&$0.898932$\\
\hline
\label{tab: JCparameters}
\end{tabular}}
\end{table}
%-------------------------------------------------------------------------
\subsection{Modified Zerilli--Armstrong model\label{sec:MZA}}
%-------------------------------------------------------------------------

As mentioned earlier, Johnson--Cook's law, widely used in the scientific community and implemented as a base in the Abaqus FEM code, is an empirical model for which the effects of strain, strain rate, and temperature on material behavior are considered separately. Since a coupling exists in reality between these three variables, it is preferable to opt for a Zerilli--Armstrong (ZA) model \cite{Hull-2011-ITD}. But the original for of ZA model has some limitations due to the fact that it is considered as two terms and to ameliorate that formulation Samantaray \eal \cite{Samantaray-2009-Thermo-viscoplastic} proposed a modified form given by the following equation:
\begin{equation}
\label{eq:MZA-model}
\sigma^y = \left(C_{1}+C_{2}\varepsilon^{p^n}\right) \exp\left[-\left(C_{3}+C_{4}\varepsilon^p\right)\left(T-T_0\right) + \left(C_{5}+C_{6}\left(T-T_0\right)\right)\ln\left( \frac{\mdot{\varepsilon}^p}{\mdot{\varepsilon}_0}\right)\right]
\end{equation}
where the $7$ coefficients $C_i$ and $n$ are the parameters of the model to be identified for a given material. The formulation of the model can be found in \cite{Samantaray-2009-Thermo-viscoplastic} and determination of its parameters is sumarized as following.

Considering the expression of the model given by Equation (\ref{eq:MZA-model}), parameter $C_1$ is deduced direcly from $\sigma^y(\varepsilon^p=0,T=T_0,\mdot{\varepsilon}^p=\mdot{\varepsilon}_0)$. When the current strain rate $\mdot{\varepsilon}^p$ is identical to the reference strain rate $\mdot{\varepsilon}_0$, the Equation (\ref{eq:MZA-model}) becomes :
\begin{equation}
\sigma^y = \left(C_{1}+C_{2}\varepsilon^{p^n}\right) \exp\left[-\left(C_{3}+C_{4}\varepsilon^p\right)\left(T-T_0\right)\right]
\end{equation}
Taking the logarithm of the previous relationship we have :
\begin{equation}
\ln\sigma^y = \ln\left(C_{1}+C_{2}\varepsilon^{p^n}\right)-\left(C_{3}+C_{4}\varepsilon^p\right)\left(T-T_0\right)
\end{equation}
By plotting the function $\ln\sigma^y\left(T-T_0\right)$ the y-intercept and the slope of this line can be deduced as $I_1=\ln\left(C_1+C_2\varepsilon^{p^n}\right)$ and $S_1=-\left(C_3+C_4\varepsilon^p\right)$.
From the expression of $I_1$ we can write:
\begin{equation}
\ln\left(\exp(I_1)-C_1\right) = \ln C_2+n\ln\varepsilon^p
\end{equation}
By plotting the function $\ln\left(\exp(I_1)-C_1\right)$ versus $\ln\varepsilon^p$ we can deduce $n$ and $C_2$. The plot of $S_1$ versus $\varepsilon^p$ allows deducing $C_3$ and $C_4$ from the y-intercept and the slope, respectively. Once the first four terms of the model are calculated, we can now take the logarithm of the entire MZA model given by Equation (\ref{eq:MZA-model}).
\begin{equation}
\ln\sigma^y = \ln\left(C_{1}+C_{2}\varepsilon^{p^n}\right) - \left(C_{3}+C_{4}\varepsilon^p\right)\left(T-T_0\right) + S_2\ln\left( \frac{\mdot{\varepsilon}^p}{\mdot{\varepsilon}_0}\right)
\end{equation}
with $S_2=C_5+C_6\left(T-T_0\right)$. Finally, the graph $\ln\sigma^y$ versus $ \ln(\mdot{\varepsilon}^p/\mdot{\varepsilon}_0)$ allows deducing $C_5$ and $C_6$ from the slope $S_2$. The parameters of MZA model are given in the Table \ref{tab: MZAparameters} while its prediction are shown in Figure \ref{fig:iCorrelationMZA}. It can be seen from those figures that the MZA model as JC model is not suitable to describe experimental at low strain rates but much better for high strain rates. The deviation between the predicted values and the experimental values is large because this model has problem to correctly describe the softening in its formulation. That is aspect of the limitations of that model as presented in \cite{tize2022generalized}.
\begin{table}[h!]
\centering{}
\caption{Parameters' constants of Modified--Zerilli-Armstrong Model}
\scalebox{0.83}{
\begin{tabular}{lllllllc}
\hline
&         &             &		   &		 &			   &&\\
Parameters&$C_1(\text{MPa})$    &$C_2(\text{MPa})$        & $C_3(\times10^{-5})$     & $C_4(\times10^{-5})$   & $C_5(\times10^{-5})$       &$C_6(\times10^{-5})$&$n$\\
&         &             &		   &	     &	           &&\\
\hline
Values&$43.0988$&$112.975$& $330.976$&$4.27581$&$14106.3$&$17.4178$&$0.305028$\\
\hline
\label{tab: MZAparameters}
\end{tabular}}
\end{table}
\begin{figure}[!ht]
\centering
\includegraphics[width=1.02\columnwidth]
{Figures/CompExpMZA}
\caption{Comparison between the experimental and predicted flow stresses by MZA model}
\label{fig:iCorrelationMZA}
\end{figure}
\FloatBarrier

@@

The Hansel-Spittel model is one of the least known models in terms of its  integration in software for finite element simulation although the determination of its parameters is easier than the JC model. A single programming command is sufficient for its identification. However, this model has some difficulties as it loses its accuracy when all parameters are taken into account. For this reason, several authors limit its use to $5$ or $6$ parameters \cite{chadha2018approach, rudnytskyj2020constitutive, el2015comparison}. In this paper the good predictions are with $5$ parameters summarized in Table \ref{tab: HSparameters}. The equation of the model is given by the following relation:

@@

\scalebox{0.9}{

@@

\end{tabular}}

@@

\scalebox{0.95}{

@@

where $A_i, B_j, C_k^l$ are the parameters (Table \ref{tab:PTM-parameters}) of the model to be identified using the procedure proposed in  \cite{tize2022generalized}. Quantities $q$, $r$, $s$ and $t$ define the degree of the polynomials used to describe the behavior of the material. The larger these quantities are, the more parameters need to be identified for the PTM model.
The determination of the parameters of this model are calculated thanks to the LMFIT Python library \cite{Newville-2016-LNL} and for more details about this model  we invite the reader to read our previous work \cite{tize2022generalized}. Thus all the parametrers of this model are calcuted with $q=5$, $r=5$, $s=1$ and $t=5$.

>>>>

%--------------------------------------------------------------------------
\section{Introduction\label{sec:Introduction}}
%--------------------------------------------------------------------------
Continuous casting and rolling mechanisms are an integral part of the technological methods of medium deformation through a continuous water-cooled crystalliser before entering a rolling mill. Some of the advantages of this method are increased metal yield, energy saving, improved quality of continuous casting and easier mechanisation and automation . Deformation being a very complex process involving both elastic and plastic deformation, is accompanied by high temperatures, high strain rates and complex and variable friction conditions \cite{He-2013}. However, a quantitative study and analysis of the deformation mechanism does not provide accurate analytical solutions . Moreover, conducting an experiment is not only time-consuming but also requires a high level of funding \cite{Changizian-2012}. To mitigate these two factors (to name but a few), the Finite Element Method (FEM) is widely used in the analysis and optimisation of process parameters of various materials . FEM effectively compensates for the shortcomings of the analytical and experimental methods and plays an important role in the correct formulation of technological parameters, the improvement of product quality and the in-depth knowledge of the deformation mechanism. For example FEM has been widely used to simulate and analyse steel hot forming processes, such as forging, rolling, extrusion and casting but the accuracy of the numerical solution provided by FEM for this type of analysis strongly depends  on the accuracy of the constitutive equation of the concerned material \cite{Qin-2010, Mandal-2009, Ji-2018}.

Indeed, hot forming processes are an essential step in the manufacturing of mechanical components and directly affect the ﬁnal quality of the products such as microstructure and mechanical properties \cite{Ashtiani-2012}. Several metallurgical phenomena such as work hardening and dynamic softening (dynamic recovery and recrystallisation), which very often occur during hot working, influence the quality of the deformation of materials. The complexity of the deformation behavior, however, leads to the fact that the deformation can only be calculated accurately by using a computer code to model the response of the material under the specified loading conditions. To this end, a mathematical formulation is often used to represent the flow behavior of materials in a form that can be used in finite element software \cite{Lin-2008-P,Wu-2012}.

Based on a thorough search of the literature many researchers have tried to predict the mechanical behavior of materials under hot working conditions using constitutive models. Therefore, various constitutive equations have been proposed to describe the flow behavior including phenomenological, physical and artiﬁcial neural network (ANN) models \cite{Lin-2011,Shin-2010,Rusinek-2010}. As for the semi-empirical models, they try to be at the border between the empirical and physical models. Among a large number of models, the Johnson--Cook (JC) \cite{Johnson-1983, Johnson-1985} and the Zerilli--Armstrong (ZA) \cite{Zerilli-1987} models are the best known and available in finite element codes. The JC model is widely used because it is simple to use and has fewer constants to be determined \cite{Khan-2004, NematNasser-2003}. However, this model, in its original state, contains some problems related to the phenomena describing the behavior of the material. We refer here to strain hardening, strain rate hardening, and thermal effects that are described separately in the JC model as analyzed by Jia \eal \cite{Jia-2021}. To overcome this difficulty, several modified forms of the Johnson--Cook model have been developed in the past \cite{Rule-1998, Vural-2003, Lin-2010, Lin-2012, Li-2013, Zhou-2019, Zhang-2015}. According to its more physical formulation, the ZA model is preferred over the JC model because it allows for the coupling between strain rate and temperature effects \cite{Johnson-1988, Voyiadjis-2005, Dey-2007}. However, the drawbacks of this model are related to the temperature limit value of the test. Indeed, this model applies only to low strain rates and for temperatures limited to $0.6~T_m$ (where $T_m$ is the melting temperature of the material) \cite{Chiou-2005, Lee-2005, Chen-2007, Lee-2006}. Moreover, this model, in its original form, does not take into account the coupling between deformation and temperature effects \cite{Samantaray-2009}. In order to reduce these shortcomings (\eg applying this model over a wider temperature range) of the ZA model, Modified Zerilli--Armstrong forms (MZA) have been proposed by several authors over the last past years \cite{NematNasser-2004, Lennon-2004, Muralli-2017, Cheng-2021, Muralli-2021}.
The hyperbolic sine Arrhenius-type constitutive model is one of the most widely used phenomenological models both for the prediction of the hot deformation behavior of materials and for the analysis of the microstructure of the studied material because   this model allows to estimate the deformation energy of the material \cite{Jonas-1969,Zhang-2012,Mostafaei-2012}. Its original formulation has been revised several times in order to properly describe the obtained experimental behavior. In order to increase the accuracy of the model, the effects of strain \cite{Slooff-2007, Li-2012, Xu-2013}, strain rate \cite{Lin-2008-C, Mandal-2009} have been considered by several researchers. For example, Sloof \eal \cite{Slooff-2007} introduced a strain-dependent parameter into the initial formulation of the Arhenius model so that their model could predict the high temperature flow behavior of a magnesium alloy. Lin \eal \cite{Lin-2008-C}, on the other hand, modified the model by introducing a strain rate compensation factor.

@@

%----------------------------------------------
\section{ Materials and Experiments \label{sec:Materials}}
%----------------------------------------------
%----------------------------------------------
\subsection{Experimental procedure}
%----------------------------------------------
The material used in this study is the modified carbon alloy AISI P20 whose chemical composition is given in the Table \ref{tab:Composition}. Cylindrical specimens were machined with an initial diameter of $10$ mm and a height of $15$ mm. Hot compression tests were conducted on a Gleeble-3800 thermo-simulator (Figure \ref{fig:Gleeble3800}), at temperatures of $1050^\circ$C,  $1050^\circ$C,  $1100^\circ$C,  $1150^\circ$C,  $1200^\circ$C and  $1250^\circ$C with strain rates of $0.001\ \text{s}^{-1}$, $0.01\ \text{s}^{-1}$, $0.1\ \text{s}^{-1}$, $1.0\ \text{s}^{-1}$, $2.0\ \text{s}^{-1}$ and $5.0\ \text{s}^{-1}$. Thin tantalum sheets were used as a lubricant material at the contact surface of anvils and specimen to minimize the frictions during the tests. The samples were heated up to temperature $1260^\circ$C  and were held for $5$ min to eliminate thermal gradients. Then cooled to test temperature where the specimen is held $1$ minute before deformation. After compression the specimen is quencted rapidly for microstructure studies. The procedure of this test is illustrated in Figure \ref{fig:Heating} where $h_0$ and $R_0$ are the height and radius before compression; $h$, $R_M$, and $R_T$ are respectively the height, big radius and small radius of the specimen after compression.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]
{Figures/Gleeble-3800}
\caption{Gleeble 3800 simulator system}
\label{fig:Gleeble3800}
\end{figure}
\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]
{Figures/GleebleDesign}
\caption{Schematic diagram of the experimental process.}
\label{fig:Heating}
\end{figure}
\begin{table}[h!]
\centering{}
\caption{Chemical Composition of the AISI P20 Steel (Weight Percent)}
\begin{tabular}{lcccccccc}
\hline
C  &Mn & Mo   & Si   & Ni       &Cr& Cu & Other& Balance\\
\hline
 0.30        &       0.89      &0.52		   &	0.34     &	 0.68          &1.86& 0.17& Microalloying& Iron\\
\hline
\label{tab:Composition}
\end{tabular}
\end{table}
%------------------------------------------------------------------------
\subsection{Flow stress behavior}
%------------------------------------------------------------------------
The flow stress curves obtained from the compression tests performed on Gleeble-3800 simulator for each test condition are shown in Figure \ref{fig:rawData}. From these curves it can be seen that the flow stress increases with increasing strain rate but decreases with increasing temperature. It should be noted that the strain influences the flow stress. Indeed, for all strain values below $0.2$, the stress increases with strain before reaching a peak value at low strain rates and then decreases to maintain a stable value until the end of the test. The peak stress value increases with strain rate but decreases with temperature. For higher strain rates ($1.0\ \text{s}^{-1}$, $2.0\ \text{s}^{-1}$ and $5.0\ \text{s}^{-1}$) the flow stress increases throughout the test. The slight increase in stress at low strain rates when the strain is large is due to friction between the specimen and the grip during the test. For low strain rates, lubrication is lost over time and therefore friction occurs. This results in a progressive increase in stress.

@@

The JC model as mentioned above, is one of the most widely used analytical models as it can be applied to several materials under different conditions of strain, strain rate and temperature . However, the formulation of this model does not take into account the simultaneous effect of strain, strain rate and temperature. Indeed, it is formulated by describing the effect of each physical parameter (strain, strain rate and temperature) separately as a factor in the mathematical expression of the model. Hence its inability to describe the phenomenon of temperature-induced softening. The equation that describes this model is given as follows \cite{Johnson-1985}:

@@

The predicted and experimental values are compared, as shown in Figure \ref{fig:iCorrelationJC}. The JC model is not much suitable to describe the ﬂow behavior of AISI P20 within entire deformation condition, and predicted values show a monotonous downward trend at various strain rates. The deviation between predicted values and experimental values is large for the low strain rates and almost acceptable at high strain rates. All parameters of JC model are in Table \ref{tab: JCparameters}.
\begin{figure}[!ht]
\centering
\includegraphics[width=1.02\columnwidth]
{Figures/CompExpJC}
\caption{Comparison between the experimental and predicted flow stresses by JC model}
\label{fig:iCorrelationJC}
\end{figure}
\begin{table}[h!]
\centering{}
\caption{Parameters' constants of Johnson-Cook Model}
\begin{tabular}{lccccc}
\hline
&         &             &		   &		  &\\
Parameters&$A(\text{MPa})$    &$B(\text{MPa})$        & $n$     & $C$   & $m$       \\
&         &             &		   &	     &	    \\
\hline
Values&$43.0988$&$114.518$& $0.313969$&$0.0995438$&$0.898932$\\
\hline
\label{tab: JCparameters}
\end{tabular}
\end{table}
%-------------------------------------------------------------------------
\subsection{Modified Zerilli--Armstrong model\label{sec:MZA}}
%-------------------------------------------------------------------------

As mentioned earlier, Johnson--Cook's law, widely used in the scientific community and implemented as a base in the Abaqus FEM code, is an empirical model for which the effects of strain, strain rate, and temperature on material behavior are considered separately. Since a coupling exists in reality between these three variables, it is preferable to opt for a Zerilli--Armstrong (ZA) model \cite{Hull-2011}. But the original for of ZA model has some limitations due to the fact that it is considered as two terms and to ameliorate that formulation Samantaray \eal \cite{Samantaray-2009} proposed a modified form given by the following equation:
\begin{equation}
\label{eq:MZA-model}
\sigma^y = \left(C_{1}+C_{2}\varepsilon^{p^n}\right) \exp\left[-\left(C_{3}+C_{4}\varepsilon^p\right)\left(T-T_0\right) + \left(C_{5}+C_{6}\left(T-T_0\right)\right)\ln\left( \frac{\mdot{\varepsilon}^p}{\mdot{\varepsilon}_0}\right)\right]
\end{equation}
where the $7$ coefficients $C_i$ and $n$ are the parameters of the model to be identified for a given material. The formulation of the model can be found in \cite{Samantaray-2009} and determination of its parameters is sumarized as following.

Considering the expression of the model given by Equation (\ref{eq:MZA-model}), parameter $C_1$ is deduced direcly from $\sigma^y(\varepsilon^p=0,T=T_0,\mdot{\varepsilon}^p=\mdot{\varepsilon}_0)$. When the current strain rate $\mdot{\varepsilon}^p$ is identical to the reference strain rate $\mdot{\varepsilon}_0$, the Equation (\ref{eq:MZA-model}) becomes :
\begin{equation}
\sigma^y = \left(C_{1}+C_{2}\varepsilon^{p^n}\right) \exp\left[-\left(C_{3}+C_{4}\varepsilon^p\right)\left(T-T_0\right)\right]
\end{equation}
Taking the logarithm of the previous relationship we have :
\begin{equation}
\ln\sigma^y = \ln\left(C_{1}+C_{2}\varepsilon^{p^n}\right)-\left(C_{3}+C_{4}\varepsilon^p\right)\left(T-T_0\right)
\end{equation}
By plotting the function $\ln\sigma^y\left(T-T_0\right)$ the y-intercept and the slope of this line can be deduced as $I_1=\ln\left(C_1+C_2\varepsilon^{p^n}\right)$ and $S_1=-\left(C_3+C_4\varepsilon^p\right)$.
From the expression of $I_1$ we can write:
\begin{equation}
\ln\left(\exp(I_1)-C_1\right) = \ln C_2+n\ln\varepsilon^p
\end{equation}
By plotting the function $\ln\left(\exp(I_1)-C_1\right)$ versus $\ln\varepsilon^p$ we can deduce $n$ and $C_2$. The plot of $S_1$ versus $\varepsilon^p$ allows deducing $C_3$ and $C_4$ from the y-intercept and the slope, respectively. Once the first four terms of the model are calculated, we can now take the logarithm of the entire MZA model given by Equation (\ref{eq:MZA-model}).
\begin{equation}
\ln\sigma^y = \ln\left(C_{1}+C_{2}\varepsilon^{p^n}\right) - \left(C_{3}+C_{4}\varepsilon^p\right)\left(T-T_0\right) + S_2\ln\left( \frac{\mdot{\varepsilon}^p}{\mdot{\varepsilon}_0}\right)
\end{equation}
with $S_2=C_5+C_6\left(T-T_0\right)$. Finally, the graph $\ln\sigma^y$ versus $ \ln(\mdot{\varepsilon}^p/\mdot{\varepsilon}_0)$ allows deducing $C_5$ and $C_6$ from the slope $S_2$. The parameters of MZA model are given in the Table \ref{tab: MZAparameters} while its prediction are shown in Figure \ref{fig:iCorrelationMZA}. It can be seen from those figures that the MZA model as JC model is not suitable to describe experimental at low strain rates but much better for high strain rates. The deviation between the predicted values and the experimental values is large because this model has problem to correctly describe the softening in its formulation. That is aspect of the limitations of that model as presented in \cite{TizeMha-2022}.
\begin{table}[h!]
\centering{}
\caption{Parameters' constants of Modified--Zerilli-Armstrong Model}
\begin{tabular}{lllllllc}
\hline
&         &             &		   &		 &			   &&\\
Parameters&$C_1(\text{MPa})$    &$C_2(\text{MPa})$        & $C_3(\times10^{-5})$     & $C_4(\times10^{-5})$   & $C_5(\times10^{-5})$       &$C_6(\times10^{-5})$&$n$\\
&         &             &		   &	     &	           &&\\
\hline
Values&$43.0988$&$112.975$& $330.976$&$4.27581$&$14106.3$&$17.4178$&$0.305028$\\
\hline
\label{tab: MZAparameters}
\end{tabular}
\end{table}
\begin{figure}[!ht]
\centering
\includegraphics[width=1.02\columnwidth]
{Figures/CompExpMZA}
\caption{Comparison between the experimental and predicted flow stresses by MZA model}
\label{fig:iCorrelationMZA}
\end{figure}
\FloatBarrier

@@

The Hansel-Spittel model is one of the least known models in terms of its  integration in software for finite element simulation although the determination of its parameters is easier than the JC model. A single programming command is sufficient for its identification. However, this model has some difficulties as it loses its accuracy when all parameters are taken into account. For this reason, several authors limit its use to $5$ or $6$ parameters \cite{Chadha-2018, Rudnytskyj-2020, Mehtedi-2015}. In this paper the good predictions are with $5$ parameters summarized in Table \ref{tab: HSparameters}. The equation of the model is given by the following relation:

@@

@@

\end{tabular}

@@

@@

where $A_i, B_j, C_k^l$ are the parameters (Table \ref{tab:PTM-parameters}) of the model to be identified using the procedure proposed in  \cite{TizeMha-2022}. Quantities $q$, $r$, $s$ and $t$ define the degree of the polynomials used to describe the behavior of the material. The larger these quantities are, the more parameters need to be identified for the PTM model.
The determination of the parameters of this model are calculated thanks to the LMFIT Python library \cite{Newville-2016} and for more details about this model  we invite the reader to read our previous work \cite{TizeMha-2022}. Thus all the parametrers of this model are calcuted with $q=5$, $r=5$, $s=1$ and $t=5$.

Comparison of predicted values of the PTM model and experimental values is shown in Figure \ref{fig:iCorrelationPTM}. The PTM model is suitable to describe the ﬂow behavior of AISI P20 steel, but for the strain rate $\mdot{\varepsilon}^p = 0.1\ \text{s}^{-1}$ the prediction in not quite good. The deviation between the predicted values and the experimental values is relatively good.
\begin{table}[!h]
\centering{}
\caption{Parameters of PTM model\label{tab:PTM-parameters}}
\begin{tabular}{cc}
\hline
&   \\
Parameters &Values\\
&   \\
\hline
$A_0 (\text{MPa})$ & $62.314$ \\
$A_1 (\text{MPa})$ & $559.827$ \\
$A_2 (\text{MPa})$ & $-2160.03$ \\
$A_3 (\text{MPa})$ & $4641.94,$ \\
$A_4 (\text{MPa})$ & $-5193.06$ \\
$A_5 (\text{MPa})$ & $2379.1$ \\
$B_0 $   & $-0.0035805$\\
$B_1 $   & $0.00282334$ \\
$B_2 $   & $-0.00775301$\\
$B_3 $   & $0.0085135$ \\
$B_4 $   & $-0.00412815$\\
$C_0^0 $ & $0.148304$ \\
$C_0^1 $ & $-0.301958$ \\
$C_0^2 $ & $2.2706$ \\
$C_0^3 $ & $-5.33003$ \\
$C_0^4 $ & $5.48971$ \\
$C_0^5 $ & $-2.11566$\\
$C_1^0 $ & $-3.74842\ \times10^{-7}$\\
$C_1^1 $ & $0.00165233$\\
$C_1^2 $ & $-0.00270485$ \\
$C_1^3 $ & $-0.000483867$\\
$C_1^4 $ & $0.00174818$ \\
\hline
\end{tabular}
\end{table}
\begin{figure}[!ht]
\centering
\includegraphics[width=1.02\columnwidth]
{Figures/CompExpPTM}
\caption{Comparison between the experimental and predicted flow stresses by PTM model }
\label{fig:iCorrelationPTM}
\end{figure}
\FloatBarrier
